---
title: "Knight Path Expectation"
author: "James"
date: "2/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Summary

I use R to simulate a knight moving randomly on a chessboard to see the average (mean) time it takes for a knight to visit each space. THe simulation predicts a mean of about 540 moves with a standard deviation of around 180 indicating that the sample mean occurs between 180 and 900 about 97% of the time. This format is a markdown file, so the code will run again as I publish it into pdf which will result in some small changes in the simulation.     

### Plan

I simulate the board with an 8x8 matrix so A1 is matrix position 1,1.  I did not feel like programming the if else tree for all legal knights moves. Instead my simulated knight picks one of its 8 moves at random, and then that candidate move is validated using acceptance/rejeciton criteria of 'being on the board'. If it is valid the knight takes that move, it is not the knight picks again. This is methodologically the same as acceptance/rejection sampling in bootstrap.

### Code

```{r}
rando_knight <- function(iterations=10){
  
  iterations <- iterations # the numer of iterations you want to see
  
  path_length <- rep(0,iterations) # initialize the vector that will hold our path lengths for each vector


for (iter in 1:iterations) {
  
  # Set Initial Postion for Knight and clear the board
  state_board = matrix(0,nrow = 8, ncol = 8)
  state_board[1,1] <- 1
  row_pos = 1
  col_pos = 1
  
  # Set Completion Status
  complete_state <- 0 # will turn to 1 once knight touches every square
  path_length_counter <- 0 
  
  # Move Around until Knight Goes to Every Spot
  while (complete_state == 0) {
    
    move_succes <- 0
    
    while (move_succes==0) {
      
      ### Move the Knight Quick to Code but needs validation (see below)
      x<-sample(1:8,1) # draw a random number between 1 and 8
      ### Forcast the knight's acording to random number.
      
      if (x==1) {
        temp_row <- row_pos+2
        temp_col <- col_pos+1
      }
      
      if (x==2) {
        temp_row <- row_pos+1
        temp_col <- col_pos+2
      }
      
      if (x==3) {
        temp_row <- row_pos-2
        temp_col <- col_pos+1
      }
      
      if (x==4) {
        temp_row <- row_pos-1
        temp_col <- col_pos+2
      }
      
      if (x==5) {
        temp_row <- row_pos+2
        temp_col <- col_pos-1
      }
      
      if (x==6) {
        temp_row <- row_pos+1
        temp_col <- col_pos-2
      }
      
      if (x==7) {
        temp_row = row_pos-2
        temp_col = col_pos-1
      }
      
      if (x==8) {
        temp_row = row_pos-1
        temp_col = col_pos-2
      }
      
      ### Validate that the move is legal aka on the boar.
      if ((temp_row <=8)&(temp_row >=1)&(temp_col <=8)&(temp_col >=1)) {
        row_pos <- temp_row
        col_pos <- temp_col
        move_succes <- 1
        path_length_counter <- path_length_counter+1
      }
      
    }
    
    ### Update Stateboard
    state_board[row_pos,col_pos] <- 1
    
    ### Check Completion COnditions
    if (sum(state_board)==64) {
      
      complete_state <- 1
      
    }
    
    
    
  }
 
  path_length[iter]<-path_length_counter
  
}
return(path_length)

}
```
### Results

Now I run the experiment by seeing how long it takes 1,000 knights to visit every spot on the board.

``` {r}
set.seed(2022) # set a seed for reproducibility.
path_lengths <- rando_knight(1000)
```

Here I publish the summary statistics

``` {r}
summary(path_lengths)

print('standard Deviation:')
sqrt(var(path_lengths))

```

Here is histogram of the results.
``` {r}
hist(path_lengths)
```